What's in a Name? Linear Temporal Logic Literally Represents Time Lines

### 2023-01-27 paper/main.tex
merge

@@

    \begin{keyword}
        Modal and Temporal Logics
        \sep Logic and Verification
        \sep Regular languages
    \end{keyword}
\end{frontmatter}
\section{Introduction}

@@

\begin{example} \label{example:air}
    Zhao and Rozier~\cite{ZR14} presents a model verification specification says ``[i]f a TSAFE command is sent to an aircraft, controller/AutoResolver should then hand off the control of this aircraft'' which corresponds to \ltl\ formula
    \begin{align*}
        \always (\texttt{tsafe.TSAFE\_command1} \land \texttt{controller.CTR\_control\_1} \\
        \limplies \nextt (\neg \texttt{controller.CTR\_control\_1}))
    \end{align*}
    For simplicity, we swap the concrete atomic proposition to $a,b$ and get
    \[
        \always (a \land b \limplies \nextt (\neg b))
    \]
    For this \ltl\ formula, our tool generates the timeline representation in \cref{fig:ex14}.
    \begin{figure}[h!]
        \centering
        \resizebox{\textwidth}{!}{
        \begin{tikzpicture}
            \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (s) {\!\!\!\!start\!\!\!\!};
            \node[shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab1) [right= of s]{\!\!\!\!$a \tand b$\!\!\!\!};
            \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab2) [above= 0.6cm of ab1]{\!\!\!\!$a \tand b$\!\!\!\!};
            \node (m) at ($(ab1)!0.5!(ab2)$) {};
            \node[shape=rectangle, fill=gray!30!white, style={minimum width=4cm, minimum height=3cm}] (g2) [right= -1cm of m] {};
            \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab1) [right= of s]{\!\!\!\!$a \tand b$\!\!\!\!};
            \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab2) [above= 0.6cm of ab1]{\!\!\!\!$a \tand b$\!\!\!\!};
            \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab3) [below= 0.6cm  of ab1]{\!\!\!\!\!\!\!\!$\tnot a\! \tor \! \tnot b$\!\!\!\!\!\!\!\!};
            \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (b1) [right= 0.75cm  of ab1]{\!\!\!\!\!\!\!\!$\tnot b$\!\!\!\!\!\!\!\!};
            \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab4) [right= .75cm of ab3]{\!\!\!\!\!\!\!\!$\dots$\!\!\!\!\!\!\!\!};
            \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab5) [right= .75cm of ab4]{\!\!\!\!\!\!\!\!$\tnot a\! \tor \! \tnot b$\!\!\!\!\!\!\!\!};
            \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab6) [right= .75cm of ab5]{\!\!\!\!\!\!\!\!$a \tand b$\!\!\!\!\!\!\!\!};
            \node[shape=rectangle, fill=gray!30!white, style={minimum width=10.9cm, minimum height=1.3cm}] (g2) [right= 0.6cm of ab6] {};
            \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab7) [right= .75cm of ab6]{\!\!\!\!\!\!\!\!$\tnot b$\!\!\!\!\!\!\!\!};
            \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab8) [right= .75cm of ab7]{\!\!\!\!\!\!\!\!$\tnot a\! \tor \! \tnot b$\!\!\!\!\!\!\!\!};
            \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab9) [right= .75cm of ab8]{\!\!\!\!\!\!\!\!$\dots$\!\!\!\!\!\!\!\!};
            \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab10) [right= .75cm of ab9]{\!\!\!\!\!\!\!\!$\tnot a\! \tor \! \tnot b$\!\!\!\!\!\!\!\!};
            \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab11) [right= .75cm of ab10]{\!\!\!\!\!\!\!\!$a \tand b$\!\!\!\!\!\!\!\!};

>>>>

\tikzset{
    partial ellipse/.style args={#1:#2:#3}{
        insert path={+ (#1:#3) arc (#1:#2:#3)}
    }
}

@@

    \begin{keyword}
        Modal and Temporal Logics
        \sep Logic and Verification
        \sep Regular languages
    \end{keyword}
\end{frontmatter}

\section{Introduction}

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
        \node[fill=white!50!gray, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (p1) {\!\!\!$p$\!\!\!};
        \node (p0) [left= 0.5cm of p1] {};
        \node (p2) [right= 0.5cm of p1] {};
        \draw[-latex, thick] (p0) edge (p1) (p1) edge (p2);

%        \draw[thick, black, -latex] (0,0) [partial ellipse=180:90:0.75cm and 0.5cm];
%        \draw[thick, black, -] (0,0) [partial ellipse=90:0:0.75cm and 0.5cm];
%
%        \draw[thick, black, -latex] (0,0) [partial ellipse=360:270:0.75cm and 0.5cm];
%        \draw[thick, black, -] (0,0) [partial ellipse=270:180:0.75cm and 0.5cm];

   \end{tikzpicture}
   ~~~~~
    \begin{tikzpicture}
        \node[fill=white!50!gray, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (p1) {\!\!\!$p$\!\!\!};
        \node (p0) [left= 0.5cm of p1] {};
        \node (p2) [right= 0.5cm of p1] {};
        \draw[-latex, thick] (p0) edge (p1) (p1) edge (p2);

%        \draw[thick, black, -latex] (0,0) [partial ellipse=180:90:0.75cm and 0.5cm];
%        \draw[thick, black, -] (0,0) [partial ellipse=90:0:0.75cm and 0.5cm];

        \draw[thick, black, -latex] (-0.15,0) [partial ellipse=360:270:1cm and 0.7cm];
        \draw[thick, black, -] (-0.15,0) [partial ellipse=270:180:1cm and 0.7cm];

  \end{tikzpicture}
  ~~~~~~
  \begin{tikzpicture}
        \node[fill=white!50!gray, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (p1) {\!\!\!$p$\!\!\!};
        \node (p0) [left= 0.5cm of p1] {};
        \node (p2) [right= 0.5cm of p1] {};
        \draw[-latex, thick] (p0) edge (p1) (p1) edge (p2);

        \draw[thick, black, -latex] (-0.05cm,0) [partial ellipse=180:90:1cm and 0.7cm];
        \draw[thick, black, -] (-0.05cm,0) [partial ellipse=90:0:1cm and 0.7cm];

        \draw[thick, black, -latex] (-0.15,0) [partial ellipse=360:270:1cm and 0.7cm];
        \draw[thick, black, -] (-0.15,0) [partial ellipse=270:180:1cm and 0.7cm];

%        \draw[thick, black, -latex] (0,0) [partial ellipse=360:270:0.75cm and 0.5cm];
%        \draw[thick, black, -] (0,0) [partial ellipse=270:180:0.75cm and 0.5cm];

  \end{tikzpicture}
\end{figure}

@@

\begin{example} \label{example:air}
    Zhao and Rozier~\cite{ZR14} presents a model verification specification says ``[i]f a TSAFE command is sent to an aircraft, controller/AutoResolver should then hand off the control of this aircraft'' which corresponds to \ltl\ formula
    \begin{align*}
        \always (\texttt{tsafe.TSAFE\_command1} \land \texttt{controller.CTR\_control\_1} \\
        \limplies \nextt (\neg \texttt{controller.CTR\_control\_1}))
    \end{align*}
    For simplicity, we swap the concrete atomic proposition to $a,b$ and get
    \[
        \always (a \land b \limplies \nextt (\neg b))
    \]
    For this \ltl\ formula, our tool generates the timeline representation in \cref{fig:ex14}.
    \begin{figure}[h!]
        \centering
        \resizebox{\textwidth}{!}{
        \begin{tikzpicture}
            \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (s) {\!\!\!\!start\!\!\!\!};
            \node[shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab1) [right= of s]{\!\!\!\!$a \tand b$\!\!\!\!};
            \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab2) [above= 0.6cm of ab1]{\!\!\!\!\!\!\!\!$\tnot a\! \tor \! \tnot b$\!\!\!\!\!\!\!\!};
            \node (m) at ($(ab1)!0.5!(ab2)$) {};
            \node[shape=rectangle, fill=gray!30!white, style={minimum width=4cm, minimum height=3cm}] (g2) [right= -1cm of m] {};
            \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab1) [right= of s]{\!\!\!\!$a \tand b$\!\!\!\!};
            \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab2) [above= 0.6cm of ab1]{\!\!\!\!\!\!\!\!$\tnot a\! \tor \! \tnot b$\!\!\!\!\!\!\!\!};
            \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab3) [below= 0.6cm  of ab1]{\!\!\!\!\!\!\!\!$\tnot a\! \tor \! \tnot b$\!\!\!\!\!\!\!\!};
            \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (b1) [right= 0.75cm  of ab1]{\!\!\!\!\!\!\!\!$\tnot b$\!\!\!\!\!\!\!\!};
            \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab4) [right= .75cm of ab3]{\!\!\!\!\!\!\!\!$\dots$\!\!\!\!\!\!\!\!};
            \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab5) [right= .75cm of ab4]{\!\!\!\!\!\!\!\!$\tnot a\! \tor \! \tnot b$\!\!\!\!\!\!\!\!};
            \node[draw, thick, fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab6) [right= .75cm of ab5]{\!\!\!\!\!\!\!\!$a \tand b$\!\!\!\!\!\!\!\!};
            \node[shape=rectangle, fill=gray!30!white, style={minimum width=10.9cm, minimum height=1.3cm}] (g2) [right= 0.6cm of ab6] {};
            \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab7) [right= .75cm of ab6]{\!\!\!\!\!\!\!\!$\tnot b$\!\!\!\!\!\!\!\!};
            \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab8) [right= .75cm of ab7]{\!\!\!\!\!\!\!\!$\tnot a\! \tor \! \tnot b$\!\!\!\!\!\!\!\!};
            \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab9) [right= .75cm of ab8]{\!\!\!\!\!\!\!\!$\dots$\!\!\!\!\!\!\!\!};
            \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab10) [right= .75cm of ab9]{\!\!\!\!\!\!\!\!$\tnot a\! \tor \! \tnot b$\!\!\!\!\!\!\!\!};
            \node[fill=white, shape=ellipse, style={minimum width=1.5cm, minimum height=1cm}] (ab11) [right= .75cm of ab10]{\!\!\!\!\!\!\!\!$a \tand b$\!\!\!\!\!\!\!\!};
